### 2018-07-10 latex/discussion.tex
edited new discussion

We here analysed the phylogenetic distribution of the three enzymes using Nam as a substrate, the two NAd-salvage pathway enzymes NADA and NamPRT as well as the Nam degrading enzyme NNMT. We found that after the first appearance of NNMT in Protostomia, a diversification of NAD-consuming reactions in Deuterostomia can be observed. Using mathematical modelling, we can explain these finding, as NNMT removes excess of Nam in cells and by that reduces product inhibition of NAD-signalling enzymes, enabeling higher fluxes through these reactions. This in turn seem to have enabled a diversification and NAD-consuming enzymes in mammals are involved in a wide variety of signalling and gene regulatory mechanisms that due to there sensitivity to NAD+, have the ability to translate differences in metabolic  states into changes in signalling and gene regulation. As NAD concentrations are consequently lowered, a high affinity enzyme for recycling of Nam is required to ensure maintenance for cellular NAD-levels for other metabolic processes. Whether it was by chance or whether an this is due to mechanistic limitations that NamPRT developed a high affinity and not NADA, or whether other environmental conditions have been responsible is impossible to deduce.  But our simulations clearly show that a high affinity of NamPRT is favourable and that it therefore appears that NNMT might have been driving NamPRT evolution. Looking at the enzyme affinities of the human enzymes it furthermore appears  that both NNMT and NamPRT reached an near optimal state, as further increase of the affinity of NamPRT and decrease of the affinity of NNMT would not result in much higher steady state NAD-concentrations or NAD-consumption fluxes, according to our simulations.  It furthermore appears that NNMT makes both NAD concentration and NAD consumption relatively independent of other processes requiring NAD, such as cell growth. In addition.

>>>>

We here omprehensively analysed the phylogenetic distribution of the three enzymes using Nam as a substrate, the two NAD-salvage pathway enzymes NADA and NamPRT as well as the Nam degrading enzyme NNMT. We found that after the first appearance of NNMT in Protostomia, a diversification of NAD-consuming reactions in Deuterostomia can be observed. By using mathematical modelling, we can explain these finding, as NNMT removes excess of Nam from cells and by that reduces product inhibition of NAD-signalling enzymes. This in turn enables higher fluxes through these reactions.  Thus the diversification of  NAD-consuming enzymes in mammals semms to have been enabled by the presence of NNMT. NAD-consuming enzymes are involved in a wide variety of signalling and gene regulatory mechanisms that due to their sensitivity to NAD+, have the ability to translate differences in metabolic  states into changes in signalling and gene regulation.  As through the removal of NAD pre-cursor NNMT, NAD concentrations are lowered, a high affinity enzyme for recycling of Nam is required to ensure maintenance for cellular NAD-levels for other metabolic processes. Whether it was by chance or whether this is due to thermodynamic or mechanistic limitations that NamPRT developed a high affinity and not NADA, or whether other environmental conditions have been responsible for the evolutionary trajectory observed, is impossible to deduce based on the chosen approach.  But our simulations clearly show that a high affinity of NamPRT is required for high NAD-consumption fluxes and NAD-concentrations and that it therefore appears that NNMT might have been driving NamPRT evolution. Looking at the enzyme affinities of the human enzymes it furthermore appears  that both NNMT and NamPRT reached an near optimal state, as further increase of the affinity of NamPRT and decrease of the affinity of NNMT would not result in much higher steady state NAD-concentrations or NAD-consumption fluxes, according to our simulations.  It furthermore appears that NNMT makes both NAD concentration and NAD consumption relatively independent of other processes requiring NAD, such as cell growth. In addition.
